Using Distributed Information Stores for Patient-Clinician Health Data Sharing

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
boo

We imagine integrated activity diaries to be used by both clinicians
and patients in several ways.  First, for nearly all types of
conditions, monitoring a person's overall daily activity and sleep
duration could easily indicate how much rest, exercise, and social
exposure they might be getting.  Moreover, longitudinal records of the
time(s) that a person gets up, eats, works, exercises and relaxes can
be analysed for irregular behaviours or habits that might be affecting
their well-being. For more specific conditions, daily
measurements of a person's blood pressure, blood oxidation, heart
rate, breathing, weight, and galvanic skin response could be useful
for tracking progress on a prescribed intervention, recovery from an
illness, or onset of a new conditions, for example, or be used to
pinpoint sources of prolonged stress.

Merely providing access to such activity records back to individuals
could have direct effects on people's behaviour, motivation and
perceived well-being.  For example, just as asking individuals to
introspect and justify their opinions and feelings has been found to
be both disruptive and reinforcing of their opinions~\cite{wilson1989introspection},
allowing people to reflect upon their
historical records of their activities might help people realise or
change their minds about how or why they feel particular ways. We feel
that there is significant room for investigation here, and potential
for research pertaining to self motivation and improvement.

@@

product & measurement indices & price range (approx. USD) \\
\hline
FitBit & general activity (walking/jogging) & \$50 \\
Nike FuelBand & general activity (walking/jogging) & \$199 \\ 
WiThings WiFi Scale & weight & \$200 \\
WiThings BPM & blood pressure & \$200 \\ 
Zeo Pro & sleep duration \& quality & \$200 \\
WakeMate & sleep duration & \$50  \\
Zensorium Tink\'{e} & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
BodyMedia CORE 2 & galvanic skin response, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
Hapifork & eating time, rate, frequency & (TBD) \\ 
\end{tabular}
\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurement sensors with automatic data uplink to cloud services }
\label{sensors}
\end{table}
}
\vspace{6cm}

@@

While few of these devices had initially been evaluated for accuracy
or certified for clinical use, increasingly device manufacturers have
begun to get independent certification. For example, BodyMedia
advertises its device as ``FDA-approved, clinically validated, most
accurate wireless activity measurement''.  A separate study of
activity-based sleep monitoring tools revealed that ``low cost
actigraphy-based approaches'' correlated well with baseline (in-lab
EEG-based measurements) for sleep duration, although less with sleep
quality metrics \cite{reid1999correlation}.

>>>>

We imagine integrated activity diaries to be used by both clinicians
and patients in several ways.  First, for nearly all types of
conditions, monitoring a person's overall daily activity and sleep
duration could easily indicate how much rest, exercise, and social
exposure they might be getting.  Moreover, longitudinal records of the
time(s) that a person gets up, eats, works, exercises and relaxes can
be analysed for irregular behaviours or habits that might be affecting
his or her well-being. For more specific conditions, daily
measurements of a person's blood pressure, blood oxidation, heart
rate, breathing, weight, and galvanic skin response could be useful
for tracking progress on a prescribed intervention, recovery from an
illness, or onset of a new conditions, for example, or be used to
pinpoint sources of prolonged stress.

Simply letting individuals see their own activity diaries could have
direct effects on people's behaviour, motivation and perceived
well-being, as well.  For example, just as asking individuals to
introspect and justify their opinions and feelings has been found to
be both disruptive and reinforcing of their
opinions~\cite{wilson1989introspection}, allowing people to reflect
upon their historical activity records might help them realise or
change their minds about how or why they feel particular ways. We feel
that there is significant room for investigation here, and potential
for research contributions pertaining to motivation and self-improvement.

@@

product & measurement indices & price range (approx. USD) \\
\hline
FitBit & general activity (walking/jogging) & \$50 \\
Nike FuelBand & general activity (walking/jogging) & \$199 \\ 
WiThings WiFi Scale & weight & \$200 \\
WiThings BPM & blood pressure & \$200 \\ 
Zeo Pro & sleep duration \& quality & \$200 \\
WakeMate & sleep duration & \$50  \\
Zensorium Tinnke & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
BodyMedia CORE 2 & galvanic skin response, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
Hapifork & eating time, rate, frequency & (TBD) \\ 
\end{tabular}
\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurement sensors with automatic data uplink to cloud services }
\label{sensors}
\end{table}
}
\vspace{6cm}

@@

While few of these devices had initially been evaluated for accuracy
or certified for clinical use, increasingly device manufacturers have
begun to get independent certification. For example, BodyMedia
advertises its device as ``FDA-approved, clincally validated, most
accurate wireless activity measurement''.  A separate study of
activity-based sleep monitoring tools revealed that ``low cost
actigraphy-based approaches'' correlated well with baseline (in-lab
EEG-based measurements) for sleep duration, although less with sleep
quality metrics \cite{reid1999correlation}.
